---
title: "Plate Reader JEC21RNAimutants 30 degrees"
author: "Liz Hughes and Thom Davies"
date: "23-02-21"
output:
  html_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Plate reader assay.

##Set up to test 2 different growth conditions in Jec21 cells compared to RNAi deletion mutants, dcR1-2 and RdRp1.
##3 Bioreps, 3 Techreps each, 3 Strains (JEC21-3084, dcR1-2-6595, RdRp1-6594) and 1 Biorep, 3 Techreps for H99.
##Grow 5ml culture from colony (1 colony = 1 Biorep)in YPD, 30C, 200 rpm ON.
##Seed at OD 600nm = 0.2; 200 ul per well.
##Run for 430 cycles.

```{r}
library("reshape2")
library(dplyr)
library(ggplot2)
library(tidyverse)
library(markdown)


```



# Read in transposed data as csv file

## ~/Documents/Edward Wallace/R/R files/Data/2021/        replace this with your path
## 20210219_JEC21RNAimut_30_TD_TRSP.csv                   this is the file name

```{r}
rawdata <- read.csv("../input/20210219_JEC21RNAimut_30_TD_TRSP.csv")

view(rawdata)
```

# Change time in seconds to time in hours

```{r}
 rawdata_hrs<-mutate(rawdata, Time = Time/3600)

```

# Tidy the data using the melt function from reshape2

```{r}
reshaped <- melt(rawdata_hrs, id=c("Time", "Temp"), variable.name="Well",
                 value.name="OD595")


summary(reshaped)
```

# Read in the Plate map data from csv file

## ~/Documents/Edward Wallace/R/R files/Data/2021/        replace this with your path
## 20210219_JEC21RNAimut_SetupCSV.csv                     this is the file name

```{r}

platemap <- read.csv("../input/20210219_JEC21RNAimut_SetupCSV.csv ")
head(platemap, n=10)


```


# Combine the reshaped data with the plate map, pairing them by Well

```{r}
annotated <- inner_join(reshaped, platemap, by="Well")
 
View(annotated)

```


# Write the annotated data set to a CSV file

## ~/Documents/Edward Wallace/Lab based work/Thom Davies project/        replace this with your path
## 20210219_JEC21RNAimut_30_annotated.csv                                this is the file name

```{r}
write.csv(annotated, "~/Documents/Edward Wallace/Lab based work/Thom Davies project/20210219_JEC21RNAimut_30_annotated.csv")
```


#Plot all ODs unnormalised

## Do this as check on the data by strain and medium

```{r plot_all_ODs}

ggplot(data=annotated, aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(Medium ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

#Plot all ODs by Medium, unnormalised

## This allows you to check the data by medium

```{r plot_all_ODs_bymedium}

ggplot(data=annotated, aes(x=Time, y=OD595, color=Medium, group=Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(Strain ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

# Which wells grow to higher OD?

## This helps you to pinpoint any outliers or odd wells with a much higher OD.

```{r find_wells_highOD}

annotated %>%
  group_by(Well) %>% 
  summarise(maxOD=max(OD595)) %>%
  arrange(desc(maxOD))
```



# Which wells have higher OD in YPD for JEC21

## You can check all the strains by changing filter(Strain=="JEC21-3084") or medium by changing filter(Medium=="YPD")

```{r find_wells_highOD YPD}

annotated %>%
  group_by(Well) %>% 
  filter(Strain=="JEC21-3084") %>%
  filter(Medium=="YPD") %>%
  summarise(maxOD=max(OD595)) %>%
  arrange(desc(maxOD))
```


## Which wells have higher OD in YNB for JEC21

```{r find_wells_highOD YNB}

annotated %>%
  group_by(Well) %>% 
  filter(Strain=="JEC21-3084") %>%
  filter(Medium=="YNB") %>%
  summarise(maxOD=max(OD595)) %>%
  arrange(desc(maxOD))
```
These are much lower than YPD, as expected.



# How stable are ODs of blank wells, to use for normalisation?

```{r plot_blank_ODs}

ggplot(data=filter(annotated,Strain==""), 
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(Medium ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

These are very stable.



# Use this to check a particular well if the blanks look a bit weird.
## We will check A6

```{r plot_A6}

ggplot(data=filter(annotated,Well=="A6"), 
      
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(Medium ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")

```
This is a single plot for well A6, YPD blank.


```{r plot_A12}

ggplot(data=filter(annotated,Well=="A12"), 
      
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(Medium ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")

```
This is a single plot for well A6, YNB blank.

# Check OD of blank wells by medium


```{r find_wells_highOD blanks}
annotated %>%
  group_by(Well, Medium) %>% 
  filter(Strain=="") %>%
  
  summarise(maxOD=max(OD595)) %>%
  arrange(desc(maxOD))
```
YPD>YNB
All the blank wells look ok to use.


## Calculate median OD for blank wells to use to normalise data.

```{r calc_blank_medians}

blank_OD_summary <- annotated %>%
  filter(Strain=="") %>%
  group_by(Medium) %>%
  summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(blank_OD_summary)
```
These all look good.



# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs)
## Normalize data

```{r subtract_blank_OD}

normalisedOD <- annotated %>%
  left_join(blank_OD_summary, by="Medium") %>%
  mutate(OD_corrected = OD595 - OD_median)

view(normalisedOD)
```



#Plot all OD corrected (or normalized)

```{r plot_all_OD_corrected}

ggplot(data=filter(normalisedOD,Strain !=""),  aes(x=Time, y=OD_corrected, color=Strain, group = Well)) + 
       geom_line() + 
       facet_grid(Medium ~ .) +
       theme_bw()+
       labs(x = "Time(Hrs)") +
       labs(y = "Absorbance 595nm")
```





```{r plot_all_OD_corrected_zoom for YNB}

ggplot(data=normalisedOD %>%
         filter(Strain != "", Medium =="YNB"), 
       aes(x=Time, y=OD_corrected, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw() +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "Growth in YNB") +
       theme(text = element_text(size = 20))+
       scale_color_brewer(palette="Set1")
```


```{r plot_all_OD_corrected_zoom for YNB wrap}

ggplot(data=normalisedOD %>%
         filter(Strain != "", Medium =="YNB"), 
       aes(x=Time, y=OD_corrected, color=Strain, group = Well)) + 
       geom_line() + 
       facet_wrap(BioRep ~ .)+
       theme_bw() +
         scale_y_continuous(trans = 'log10')+
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "Growth in YNB") +
       theme(text = element_text(size = 20))+
       scale_color_brewer(palette="Set1")
```



# Plot for H99 only

```{r plot_all_OD_corrected for H99 only wrap}

ggplot(data=normalisedOD %>%
         filter(Strain == "H99"), 
         aes(Time, OD_corrected, group = Well, colour = Medium)) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
         facet_wrap(BioRep ~ .) +
         theme_bw() +
         scale_y_continuous(limits=c(0,2),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "H99") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")


   
```


# Plot for JEC21-3084 only

```{r plot_all_OD_corrected for JEC21-3084 only wrap}

ggplot(data=normalisedOD %>%
         filter(Strain == "JEC21-3084"), 
         aes(Time, OD_corrected, group = Well, colour = Medium)) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
         facet_wrap(BioRep ~ .) +
         theme_bw() +
         scale_y_continuous(limits=c(0,2),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "JEC21-3084") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")

```

# Plot for dcR1-2-6595 only

```{r plot_all_OD_corrected for dcR1-2-6595 only}
ggplot(data=normalisedOD %>%
         filter(Strain == "dcR1-2-6595"), 
         aes(Time, OD_corrected, group = Well, colour = Medium)) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
         facet_wrap(BioRep ~ .) +
         theme_bw() +
         scale_y_continuous(limits=c(0,2),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "dcR1-2-6595") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")
```

# Plot for RdRp1-6594 only
```{r plot_all_OD_corrected for RdRp1-6594}
ggplot(data=normalisedOD %>%
         filter(Strain == "RdRp1-6594"), 
         aes(Time, OD_corrected, group = Well, colour = Medium)) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
         facet_wrap(BioRep ~ .) +
         theme_bw() +
         scale_y_continuous(limits=c(0,2),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "RdRp1-6594") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")
```



# Plot for All strains and media by BioRep

```{r plot_all_OD_corrected for all wrap}

ggplot(data=normalisedOD %>%
         filter(Strain != ""),
       aes(Time, OD_corrected, colour = Strain, group = Well)) + 
       geom_line() + 
       geom_hline(aes(yintercept=0.5), linetype="dotted", colour="black") +
       facet_wrap(BioRep ~ .) +
       theme_bw() +
        scale_y_continuous(limits=c(0,1.5),expand=c(0,0)) +
        labs(x = "Time(Hrs)") +
        labs(y = "Absorbance 595nm") +
      theme(text = element_text(size = 10))+
      scale_color_brewer(palette="Set1")

```




