---
title: "Plate Reader JEC21RNAimutants 8Day culturability assay 30 degrees"
author: "Liz Hughes and Thom Davies"
date: "12-03-21"
output:
  html_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Plate reader assay.

##Set up to test growth of Jec21 cells compared to RNAi deletion mutants, dcR1-2 and RdRp1. 1 colony for each biorep into 5 ml YPD ON (STAT1), 1:100 dilution of this cututre into 60ml YPD ON (STAT2), 25ml split into 2 flask; 1 incubated at 60rpm, 30 degrees (Hypoxia) and the other incubated at 180rpm, 30 degrees (Normoxia) for 8 days.
## Use Cfu to count number of cells /ml. Aliquot approx 100 cells/well. Grow at 30 degrees for 5 more days. Take OD 595.
##2 Bioreps, 5 Techreps each, 3 Strains (JEC21-3084, dcR1-2-6595, RdRp1-6594).


```{r}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(markdown)


```
# Read in transposed data as csv file

```{r}
rawdata <- read.csv("/Users/s2124188/Documents/Physical Rotations/Wallace & Bayne/CFU Assays/OD v Cell No/20210310_All2_csv.csv")

view(rawdata)
```

# Separate Strain into Strain, BioRep and Condition

```{r separate conditions}
Data <- separate(rawdata, 
         Strain, 
         into = c("Strain" , "BioRep" , "Condition"), 
         sep = "_", 
         remove = TRUE)
view(Data)
```


## Calculate median OD for blank wells to use to normalise data.

```{r calc_blank_medians}

blank_OD_summary <- Data %>%
  filter(Strain=="blank") %>%
  group_by(Condition) %>%
    summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(blank_OD_summary)
```

# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs)
## Normalize data

```{r subtract_blank_OD}

normOD <- Data %>%
  left_join(blank_OD_summary, by="Condition") %>%
  mutate(OD_corrected = OD595 - OD_median)
view(normOD)
```


#Create subset dcR
```{r dcR}

dcR <- normOD %>%
filter(Strain=="dcR")
```


#Plot dcR, normalised
```{r plot_dcR}

ggplot(data=dcR, aes(x=Cell_No, y=OD595, color=BioRep)) + 
         geom_boxplot() + 
       theme_bw()+
 facet_wrap(Condition ~.)+
   labs(x = "Cell_No") +
   labs(y = "Absorbance 595nm") +
   labs(title = "dcR") +
  theme(text = element_text(size = 20))+
   scale_color_brewer(palette="Set1")
```

#Create subset RdRp
```{r RdRp}

RdRp <- normOD %>%
filter(Strain=="RdRp")
```


#Plot RdRp, normalised
```{r plot_RdRp}

ggplot(data=RdRp, aes(x=Cell_No, y=OD595, color=BioRep)) + 
         geom_boxplot() + 
       theme_bw()+
 facet_wrap(Condition ~.)+
   labs(x = "Cell_No") +
   labs(y = "Absorbance 595nm") +
   labs(title = "RdRp") +
  theme(text = element_text(size = 20))+
   scale_color_brewer(palette="Set1")
```

#Create subset JEC21
```{r JEC21}

JEC21 <- normOD %>%
filter(Strain=="JEC21")
```

#Plot JEC21, normalised
```{r plot_JEC21}

ggplot(data=JEC21, aes(x=Cell_No, y=OD595, color=BioRep)) + 
         geom_boxplot() + 
       theme_bw()+
 facet_wrap(Condition ~.)+
  labs(x = "Cell_No") +
   labs(y = "Absorbance 595nm") +
   labs(title = "JEC21") +
  theme(text = element_text(size = 20))+
   scale_color_brewer(palette="Set1")
```










